---
title: "pca_human_nini's_version"
author: "Nini"
date: "2023-08-04"
output: 
  pdf_document: default
  html_document: default
---

``````{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(factoextra)
library(dplyr)
library(tidyverse)
library(DESeq2)
options(ggrepel.max.overlaps = 50)
```


```{r functions, echo=FALSE}
pca_analysis <- function(data, labels, header = "title") {
res.pca <- prcomp(t(data))


a <- fviz_eig(res.pca, title = header)


b <- fviz_pca_ind(res.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE,     # Avoid text overlapping,
             title = header
)


c <- fviz_pca_var(res.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE,     # Avoid text overlapping
             title = header
)

#fviz_pca_biplot(res.pca, repel = TRUE,
#                col.ind = "cos2",  # Individuals color
#                gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
#                col.var = "#2E9FDF" # Variables color
#)

groups <- as.factor(paste(labels$V2, labels$V3))
d <- fviz_pca_ind(res.pca,
             col.ind = groups, # color by groups
             #palette = c("#00AFBB",  "#FC4E07"),
             #addEllipses = TRUE, # Concentration ellipses
             #ellipse.type = "confidence",
             legend.title = "Groups",
             col.var = "#2E9FDF", # Variables color
             repel = TRUE,
             title = header
)
plot_list <- list(a,b,c,d)
return(plot_list)
}

```

## NOER and ER



```{r data, echo=FALSE}
getwd()
data_star <- as.matrix(read.table(file = "/Users/nini/Documents/EBV/human/gene.counts.star 2", row.names = 1, header = TRUE))
head(data_star)
data_star_nuc <- as.data.frame(data_star) %>% select("MUC5708_S14.tr1", "MUC5708_S14.tr2", "MUC5708_S14.tr3", "SAR59.6.tr1", "SAR59.6.tr2", "SAR59.10.tr1", "SAR59.10.tr2", "MUC5706_S12.tr1", "MUC5706_S12.tr2", "MUC5706_S12.tr3", "SAR59.5.tr1", "SAR59.5.tr2", "SAR59.9.tr1", "SAR59.9.tr2")


data_star_cyto <- as.data.frame(data_star) %>% select("MUC5707_S13.tr1", "MUC5707_S13.tr2", "MUC5707_S13.tr3", "MUC3775_S61.tr1", "MUC3775_S61.tr2", "MUC3775_S61.tr3", "MUC3776_S62.tr1", "MUC3776_S62.tr2", "MUC3776_S62.tr3", "MUC5705_S11.tr1", "MUC5705_S11.tr2", "MUC5705_S11.tr3", "MUC3772_S58.tr1", "MUC3772_S58.tr2", "MUC3772_S58.tr3", "MUC3773_S59.tr1", "MUC3773_S59.tr2", "MUC3773_S59.tr3")

data_star_chx_nuc <- as.data.frame(data_star) %>% select("SAR59.11.tr1","SAR59.11.tr2","SAR59.12.tr1","SAR59.12.tr2","MUC5702_S8.tr1","MUC5702_S8.tr2","MUC5702_S8.tr3","MUC5704_S10.tr1","MUC5704_S10.tr2","MUC5704_S10.tr3","SAR59.7.tr1","SAR59.7.tr2","SAR59.8.tr1","SAR59.8.tr2")


data_star_chx_cyto <- as.data.frame(data_star) %>% select("MUC5701_S7.tr1","MUC5701_S7.tr2","MUC5701_S7.tr3","MUC5703_S9.tr1","MUC5703_S9.tr2","MUC5703_S9.tr3","MUC3778_S64.tr1","MUC3778_S64.tr2","MUC3778_S64.tr3","MUC3779_S65.tr1","MUC3779_S65.tr2","MUC3779_S65.tr3","MUC3781_S67.tr1","MUC3781_S67.tr2","MUC3781_S67.tr3","MUC3782_S68.tr1","MUC3782_S68.tr2","MUC3782_S68.tr3")

data_star_NOER_ER <- as.data.frame (data_star) %>%
select("MUC5708_S14.tr1", "MUC5708_S14.tr2", "MUC5708_S14.tr3", "SAR59.6.tr1", "SAR59.6.tr2", "SAR59.10.tr1", "SAR59.10.tr2", "MUC5706_S12.tr1", "MUC5706_S12.tr2", "MUC5706_S12.tr3", "SAR59.5.tr1", "SAR59.5.tr2", "SAR59.9.tr1", "SAR59.9.tr2","MUC5707_S13.tr1", "MUC5707_S13.tr2", "MUC5707_S13.tr3", "MUC3775_S61.tr1", "MUC3775_S61.tr2", "MUC3775_S61.tr3", "MUC3776_S62.tr1", "MUC3776_S62.tr2", "MUC3776_S62.tr3", "MUC5705_S11.tr1", "MUC5705_S11.tr2", "MUC5705_S11.tr3", "MUC3772_S58.tr1", "MUC3772_S58.tr2", "MUC3772_S58.tr3", "MUC3773_S59.tr1", "MUC3773_S59.tr2", "MUC3773_S59.tr3")

data_star_CHX_all <- as.data.frame (data_star) %>%
select("MUC5701_S7.tr1","MUC5701_S7.tr2","MUC5701_S7.tr3","MUC5703_S9.tr1","MUC5703_S9.tr2","MUC5703_S9.tr3","MUC3778_S64.tr1","MUC3778_S64.tr2","MUC3778_S64.tr3","MUC3779_S65.tr1","MUC3779_S65.tr2","MUC3779_S65.tr3","MUC3781_S67.tr1","MUC3781_S67.tr2","MUC3781_S67.tr3","MUC3782_S68.tr1","MUC3782_S68.tr2","MUC3782_S68.tr3","SAR59.11.tr1","SAR59.11.tr2","SAR59.12.tr1","SAR59.12.tr2","MUC5702_S8.tr1","MUC5702_S8.tr2","MUC5702_S8.tr3","MUC5704_S10.tr1","MUC5704_S10.tr2","MUC5704_S10.tr3","SAR59.7.tr1","SAR59.7.tr2","SAR59.8.tr1","SAR59.8.tr2")


```

``````{r nucleus, echo=FALSE}
labels <- data.frame(read.csv(file = "/Users/nini/Documents/EBV/human/er-noer.all.labels.map", header = FALSE, sep = "\t"))

data_star <- data_star_nuc
star <- pca_analysis(data_star, labels, "not normalized")

data_star_rlog <- rlogTransformation(as.matrix(data_star))
data_star_vst <- varianceStabilizingTransformation(as.matrix(data_star))

rownames(data_star_rlog) <- rownames(data_star)
rownames(data_star_vst) <- rownames(data_star)

star_rlog <- pca_analysis(data_star_rlog, labels, "rlog transform")
star_vst <- pca_analysis(data_star_vst, labels, "vst transform")

star[1]
star_rlog[1]
star_vst[1]

star[2]
star_rlog[2]
star_vst[2]

star[3]
star_rlog[3]
star_vst[3]

star[4]
star_rlog[4]
star_vst[4]

```
#chx_nucl

``````{r chx_nucleus, echo=FALSE}
labels <- data.frame(read.csv(file = "/Users/nini/Documents/EBV/human/chx.nucl.labels.map", header = FALSE, sep = "\t"))

data_star <- data_star_chx_nuc
star <- pca_analysis(data_star, labels, "not normalized")

data_star_rlog <- rlogTransformation(as.matrix(data_star))
data_star_vst <- varianceStabilizingTransformation(as.matrix(data_star))

rownames(data_star_rlog) <- rownames(data_star)
rownames(data_star_vst) <- rownames(data_star)

star_rlog <- pca_analysis(data_star_rlog, labels, "rlog transform")
star_vst <- pca_analysis(data_star_vst, labels, "vst transform")

star[1]
star_rlog[1]
star_vst[1]

star[2]
star_rlog[2]
star_vst[2]

star[3]
star_rlog[3]
star_vst[3]

star[4]
star_rlog[4]
star_vst[4]

```

## Cyto samples

```{r cyto, echo=FALSE}
data_star <- data_star_cyto

labels <- data.frame(read.csv(file = "./er-noer.cyto.labels.map", header = FALSE, sep = "\t"))

star <- pca_analysis(data_star, labels, "not normalized")

data_star_rlog <- rlogTransformation(as.matrix(data_star))
data_star_vst <- varianceStabilizingTransformation(as.matrix(data_star))

rownames(data_star_rlog) <- rownames(data_star)
rownames(data_star_vst) <- rownames(data_star)

star_rlog <- pca_analysis(data_star_rlog, labels, "rlog transform")
star_vst <- pca_analysis(data_star_vst, labels, "vst transform")


star[1]
star_rlog[1]
star_vst[1]

star[2]
star_rlog[2]
star_vst[2]

star[3]
star_rlog[3]
star_vst[3]

star[4]
star_rlog[4]
star_vst[4]

```


## ChX-Cyto samples

```{r chx-cyto, echo=FALSE}
data_star <- data_star_chx_cyto

labels <- data.frame(read.csv(file = "/Users/nini/Documents/EBV/human/chx.cyto.labels.map", header = FALSE, sep = "\t"))

star <- pca_analysis(data_star, labels, "not normalized")

data_star_rlog <- rlogTransformation(as.matrix(data_star))
data_star_vst <- varianceStabilizingTransformation(as.matrix(data_star))

rownames(data_star_rlog) <- rownames(data_star)
rownames(data_star_vst) <- rownames(data_star)

star_rlog <- pca_analysis(data_star_rlog, labels, "rlog transform")
star_vst <- pca_analysis(data_star_vst, labels, "vst transform")


star[1]
star_rlog[1]
star_vst[1]

star[2]
star_rlog[2]
star_vst[2]

star[3]
star_rlog[3]
star_vst[3]

star[4]
star_rlog[4]
star_vst[4]

```


```{r all data, echo=FALSE}

data_star <- as.matrix(read.table(file = "/Users/nini/Documents/EBV/human/gene.counts.star", row.names = 1, header = TRUE))
labels <- data.frame(read.csv(file = "/Users/nini/Documents/EBV/all.labels.map", header = FALSE, sep = "\t"))




star <- pca_analysis(data_star, labels, "not normalized")

data_star_rlog <- rlogTransformation(as.matrix(data_star))
data_star_vst <- varianceStabilizingTransformation(as.matrix(data_star))

rownames(data_star_rlog) <- rownames(data_star)
rownames(data_star_vst) <- rownames(data_star)

star_rlog <- pca_analysis(data_star_rlog, labels, "rlog transform")
star_vst <- pca_analysis(data_star_vst, labels, "vst transform")


star[1]
star_rlog[1]
star_vst[1]

star[2]
star_rlog[2]
star_vst[2]

star[3]
star_rlog[3]
star_vst[3]

star[4]
star_rlog[4]
star_vst[4]

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
